---
title: "C15311966 CA notebook"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
  output: html_notebook
---

## Setup
```{r}

needed_packages <- c("pastecs","skimr", "ggplot2", "psych", "semTools", "FSA", "car", "coin", "rstatix", "effectsize", "tidyverse")                      
# Extract not installed packages
not_installed <- needed_packages[!(needed_packages %in% installed.packages()[ , "Package"])]    
# Install not installed packages
if(length(not_installed)) install.packages(not_installed)


library(tidyverse) # useful tooling
library(skimr) # for descriptions
library(ggplot2) # plots
library(pastecs) # for descriptive statistic summaries
library(psych) # useful descriptive functions
library(semTools) #For skewness and kurtosis
library(FSA) #For percentage
library(car) # For Levene's test for homogeneity of variance 
library(effectsize) #To calculate effect size for t-test

student_pref = read_csv("sperformance-dataset.csv")
getmode <- function(v) {
        uniqv <- unique(v)
        uniqv[which.max(tabulate(match(v, uniqv)))]
}
# Hypothesis Question: Does a student’s parents’ level of education influence their level of academic achievement?

```

```{r Generating descriptive statistics}
# fill with the rest of summary statistics
# find kurtosis function

getmode(student_pref$mG3)

getmode(student_pref$Medu)

# just checking variance with function
var(student_pref$Medu)
var(student_pref$mG3)

# median for final maths grades
median(student_pref$mG3)
# median for mothers education levels
median(student_pref$Medu)

# IQR for final maths grades
quantile(student_pref$mG3) # Q3 (14) - Q1 (8) = 6
# IQR for mothers education levels
quantile(student_pref$Medu) # Q3 (4) - Q1 (2) = 2



skim(student_pref)
skim(student_pref$Medu)
skim(student_pref$Fedu)


# wants to pursue higher education maths
# No - 18 Yes - 364
table(student_pref$higher.m)
# No - 18 Yes - 364
table(student_pref$higher.p)
# very few students dont want to pursue higher education 

```

## H0: There is no difference in students scores in regards to mothers education level.
## HA: There is a difference in students scores in regards to mothers education level.
### tlifesat Examine/Assess Normality
```{r mothers education assessing normality}
gg = ggplot(student_pref, aes(x=Medu))
gg = gg + geom_histogram(bins = 5, colour="cornflowerblue", aes(y=..density.., fill=..count..))
gg = gg + labs(x="Mothers education levels")
#adding a normal curve
#use stat_function to compute a normalized score for each value of Medu
#pass the mean and standard deviation
#use the na.rm parameter to say how missing values are handled
gg = gg + stat_function(fun=dnorm, color="red",args=list(mean=mean(student_pref$Medu, na.rm=TRUE), sd=sd(student_pref$Medu, na.rm=TRUE)))
#to display the graph 
gg

#Create a qqplot for mothers education levels
qqnorm(student_pref$Medu, main = "Mothers education levels")
qqline(student_pref$Medu, col = "blue", lwd = 2)

#stat.desc is a function from pastecs - make sure you include the basic switch=F to ensure you don't get scientific notation
pastecs::stat.desc(student_pref$Medu, basic=F)
dlife = pastecs::stat.desc(student_pref$Medu, basic=F)

#We can make our decision based on the value of the standardized score for skew and kurtosis
#We divide the skew statistic by the standard error to get the standardized score
#This will indicate if we have a problem
edu_skew = semTools::skew(student_pref$Medu)
edu_kurt = semTools::kurtosis(student_pref$Medu)
# skew falls slightly above +/-3 
edu_skew[1]/edu_skew[2]

# kurtosis falls inline with +/-10
edu_kurt[1]/edu_kurt[2]

#and by calculating the percentage of standardized scores for the variable itself that are outside our acceptable range
#This will tell us how big a problem we have
# Calculate the percentage of standardized scores that are greater than 1.96
# the perc function which is part of the FSA package which calculate the percentage that are within a range - you can look for greater than "gt", greater than or equal "geq", "gt", less than or equal "leq",  or less than "lt"),
# scale is a function that creates z scores, abs gets absolute value

mothers_education_level = abs(scale(student_pref$Medu))

FSA::perc(as.numeric(mothers_education_level), 1.96, "gt")
FSA::perc(as.numeric(mothers_education_level), 3.29, "gt")

```
```
Mothers education can be considered to follow a normal distribution, 100% of standardized scores falling within +/- 3.29). Therefore a parametric difference test.
```


```{r student final grade assessing normality}
gg = ggplot(student_pref, aes(x=mG3))
gg = gg + geom_histogram(bins =6, colour="cornflowerblue", aes(y=..density.., fill=..count..))
gg = gg + labs(x="Final student grades")
#adding a normal curve
gg = gg + stat_function(fun=dnorm, color="red", args=list(mean=mean(student_pref$mG3, na.rm=TRUE), sd=sd(student_pref$mG3, na.rm=TRUE)))
#to display the graph 
gg

#Create a qqplot for Final student grades
qqnorm(student_pref$mG3, main = "Final student grades")
qqline(student_pref$mG3, col = "blue", lwd = 2)

pastecs::stat.desc(student_pref$mG3, basic=F)
dlife = pastecs::stat.desc(student_pref$mG3, basic=F)


final_grade_skew = semTools::skew(student_pref$mG3)
efinal_grade_kurt = semTools::kurtosis(student_pref$mG3)
# skew falls above +/-3 
final_grade_skew[1]/final_grade_skew[2]

# kurtosis falls inline with +/-10
efinal_grade_kurt[1]/efinal_grade_kurt[2]


student_grades = abs(scale(student_pref$mG3))

FSA::perc(as.numeric(student_grades), 1.96, "gt")
FSA::perc(as.numeric(student_grades), 3.29, "gt")

```

```
Students final grades can be considered to follow a normal distribution, 100% of standardized scores falling within +/- 3.29). Therefore a parametric difference test.
```


### Differences - Parametric Tests
#### Independent t-test
##### Describe Life Satisfaction by group (have children v do not have children - variable child)
```{r H-Tests}
# Get descriptive statistics by group - output as a matrix
# psych::describeBy(student_pref$Medu, student_pref$mG3, mat=TRUE)
# Use 1 way ANOVA

# get groups and plot them with qqplot
# mG3 scores and mothers education 
# split mG3 into 20 groups ?

#qqnorm(student_pref$, main = "")
#qqline(student_pref$, col = "blue", lwd = 2)





```
Apply your understanding to the dataset provided**<br>
Provide an example of each general type of descriptive and visual for variables from the dataset
Assess at least one variable for normality

Identify some at least one correlation between your concepts of interest you want to investigate
State a hypothesis

Using your previous assessment of the data and conducting any additional assessment required, identify the most appropriate test to use
Conduct the test
Interpret the output

Report the output in text adhering to APA guidelines
Identify at least one difference effect involving a concept of interest and involving two groups you want to investigate
State a hypothesis

Using your previous assessment of the data and conducting any additional assessment required, identify the most appropriate test to use
Conduct the test
Interpret the output

Report the output in text adhering to APA guidelines
Identify at least one difference effect involving a concept of interest and involving more than two groups you want to investigate
State a hypothesis

Using your previous assessment of the data and conducting any additional assessment required, identify the most appropriate test to use
Conduct the test
Interpret the output

Report the output in text adhering to APA guidelines


```
### 

```

```

### 

```

```

